\documentclass[twocolumn]{article}

\usepackage{abstract}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.6in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{mathtools} % \multlined
\usepackage{siunitx}  % use \si{\angstrom} for Angstrom

\usepackage{subfig}
\usepackage{xspace}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\newcommand{\pygbe}{\texttt{PyGBe}\xspace}
\newcommand{\gmres}{\textsc{gmres}\xspace}
\newcommand{\bem}{\textsc{bem}\xspace}
\newcommand{\fmm}{\textsc{fmm}\xspace}
\newcommand{\kifmm}{\textsc{kifmm}\xspace}
\newcommand{\ncrit}{n_{\mathrm{crit}}}  % number of particles per leaf
\newcommand{\ses}{\textsc{ses}\xspace}
\newcommand{\msms}{\texttt{\textsc{msms}}\xspace}
\newcommand{\ie}{\textit{i}.\textit{e}., }

\graphicspath{{figs/}}

\title{High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-Exafmm}

\author{%
    Tingyu Wang\thanks{Department of Mechanical and Aerospace Engineering, The George Washington University, Washington, DC, USA}%
    \and Christopher D. Cooper\thanks{Department of Mechanical Engineering and Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso, Universidad T\'ecnica Federico Santa Mar\'ia, Valpara\'iso, Chile}%
    \and Timo Betcke\thanks{Department of Mathematics, University College London, UK}%
    \and Lorena A. Barba\samethanks[1]%
}

\date{}

\begin{document}

\twocolumn[
\maketitle

%% abstract
\begin{onecolabstract}
Biomolecular electrostatics is key in protein function and the chemical processes affecting it.
Implicit-solvent models expressed by the Poisson-Boltzmann (PB) equation can provide insights with less computational power than full atomistic models, making large-system studies---at the scale of viruses, for example---accessible to more researchers.
    This paper presents a high-productivity and high-performance PB solver based on Exafmm, a fast multipole method (FMM) library, and Bempp, a Galerkin boundary element method (BEM) package.
    Bempp-Exafmm tightly integrates an easy-to-use Python interface with well-optimized computational kernels that are written in compiled languages.
    Thanks to Python's rich ecosystem in scientific computing, users can perform PB simulations interactively via Jupyter notebooks, which opens up the possibility for faster prototyping and analyzing.
    We provide results showcasing the capability of our software, confirming correctness, and evaluating its performance with problem sizes between 8,000 and 2 million boundary elements.
    A small study comparing two variants of the boundary integral formulation in regards to algebraic conditioning showcases the power of this interactive computing platform to give useful answers with just a few lines of code.
    As a form of solution verification, mesh refinement studies with a spherical geometry as well as with a real biological structure (5PTI) confirm convergence at the expected $1/N$ rate, for $N$ boundary elements.
    Performance results include timings, breakdowns, and computational complexity.
    Exafmm offers evaluation speeds of just a few seconds for tens of millions of points, and $\mathcal{O}(N)$ scaling.
    This allowed computing the solvation free energy of a Zika virus, represented by 1.6 million atoms and 10 million boundary elements, at 80-min runtime on a single compute node (dual 20-core Intel Xeon Gold 6148).
    All results in the paper are presented with utmost care for reproducibility. 
    Readers can find input data (meshes and pqr files) in a Zenodo archive, Jupyter notebooks and secondary data necessary to reproduce all the figures in the paper's GitHub repository at \url{https://github.com/barbagroup/bempp_exafmm_paper/}, and all the software in version-controlled repositories under permissive standard licenses.

\end{onecolabstract}
]
%% keyword
%\begin{keyword}
%    boundary integral equation \sep boundary element method \sep Galerkin method \sep fast multipole method \sep
%    Python \sep biomolecular electrostatics \sep implicit solvent \sep Poisson-Boltzmann \sep solvation free energy
%\end{keyword}

% body of paper
\section{Introduction}\label{sec:intro}
\input{introduction}

\section{Results}\label{sec:results}
\input{results}

\section{Discussion} \label{sec:discussion}
\input{discussion}

\small{
\section{Methods}\label{sec:methods}
\input{methods_formulation}
\input{methods_bempp}
\input{methods_exafmm}
}

\section{Data availability}
We deposited the meshes and \texttt{pqr} files on the Zenodo service: \href{http://doi.org/10.5281/zenodo.4568768}{doi:10.5281/zenodo.4568768}.
The raw and secondary data for all results are available in the archival deposit of our paper’s GitHub repository: \href{http://doi.org/10.5281/zenodo.4568951}{doi:10.5281/zenodo.4568951}.

\section{Code availability}
Exafmm is available at \href{https://github.com/exafmm/exafmm-t}{https://github.com/exafmm/exafmm-t}.
Bempp-cl is available at \href{https://github.com/bempp/bempp-cl}{https://github.com/bempp/bempp-cl}.
The scripts for plotting and rerunning our experiments are available in the archival deposit of our paper’s GitHub repository: \href{http://doi.org/10.5281/zenodo.4568951}{doi:10.5281/zenodo.4568951}.

\bibliography{./reference}{}
\bibliographystyle{elsarticle-num}

\section*{Acknowledgments}
We thank Dr. Sergio Pantano for providing us with the parameterized structure of the Zika virus capsid.
CDC acknowledges support by ANID (Agencia Nacional de Investigaci\'{o}n y Desarrollo) through PIA/APOYO AFB180002.
TB was supported by Engineering and Physical Sciences Research Council Grant EP/V001531/1.
LAB acknowledges funding from the National Science Foundation via award \#1747669.

\section*{Author contributions}
LAB and TB conceived this project. 
TW wrote the version of the Exafmm code used in this work, the Python bindings and the Bempp integration. 
TB gave technical support on Bempp usage and wrote code to aid the integration.
CDC gave conceptual advice and helped set up computational experiments.
TW ran the calculations and prepared the figures.
TW, CDC, TB, and LAB discussed and guided the conduct of the research, and interpreted the results.
TW wrote the first draft of the manuscript, and all authors contributed materially to the writing and revising.
LAB guided the data management and guarantees the preservation of the full research compendium for this work. 
All authors confirm that the figures and conclusions accurately reflect the research.


\section*{Competing interests}
The authors declare no competing interests.

\end{document}
